%This project documents the development of a smartphone application for the Android platform. The application, UniNav, is an indoor navigation mobile application primarily designed for the students at the University of Aalborg and the layout of the Cassiopeia building to help them in checking their agenda for the lectures they are subscribed to, and find the direction where a specific lecture is held or a specific point of interest within the building. It also helps the different visitors of the University in finding of some specific point of interests inside the building.\\

%The development of the implemented application has been done by using Android platform with eclipse IDE as programming environment and Java as programming language. As development methodology, an   agile development method, scrum supported by pair programming, has been slightly adapted to accommodate the project environment. The application provides the map services functionalities implemented by using SmartCampusAAU library, which used Wi-Fi fingerprinting technologies. The scheduling information will come from a remote web application i.e. Moodle in our case, fetched into the Android application, which has an integrated database management system, SQLite which will handle these data.\\ A usability evaluation has been conducted to verify that the developed application meets the needs of the user. The analysis of the results from the usability evaluation helped us to strengthen the performances of the UniNav. The perspectives and suggestions on the developed application, ending by an overall conclusion have been discussed.


% This report documents the development of UniNav, a smartphone application primarily designed to help university students in keeping up to date with their agenda, as well as providing indoor navigation for locating lectures and other points of interests. It also features a visitor mode, helping visitors find points of interests inside buildings.

% The development consists of two parts, a mobile app for the Android platform, and a web service, with most focus on the mobile application. Currently, the app is designed to work with the Cassiopeia building of Aalborg University, and the software engineer study. The web service automatically handles fetching calendar updates from the internal system, Moodle, and the app is used for presenting this and providing indoor navigation to points of interests, for example where a lecture is held. The report also documents how we used the Scrum development methodology for conducting the development. 

Java is one of the most known and widely used programming languages and several attempts have been made to bring Java into the world of Real-Time systems. Recently, Java has also been targeted the development of Safety-Critical systems with the new standard, JSR-302 Safety-Critical Java (SCJ), to facilitate the development of systems, certifiable with standards such as DO-178B Level A. The SCJ specification is still in draft, and lacks implementations and use cases in order to be thoroughly evaluated. This report presents a study of SCJ and the application of this for two use cases.
We examine the draft version of the SCJ specification and an implementation of SCJ on the Java Optimized Processor (JOP). We take the specification and its implementation on JOP and apply this in the development of safety-critical software in collaboration with GomSpace that specialises in cubesat and nano-satellite platforms. We develop an SCJ version of their Cubesat Space Protocol (CSP), a network-layer delivery protocol for cubesats. We take this implementation and develop a Watchdog module in SCJ that utilises the CSP protocol to monitor other modules in a safety-critical environment and analyse the applicability of SCJ from a theoretical point of view with regard to theory in real-time software such as response-time analysis of applications.
The result of this study is an analysis and findings on the applicability of SCJ in its current state, for the development of safety-critical systems with emphasis on transitioning ones Java experience to SCJ.